---
title: "Arbeidskrav 2 - Regresjonsmodeller"
format: html
bibliography: bib.bib
author: Eskil Strand
editor_options: 
  chunk_output_type: console
---



# Part 1 - Lactate thresholds

There are several suggestions on how to best capture the physiological "essence" of the lactate threshold test [See @RN2511, Chapter 6]. A simple, and very common way to analyze the relationship between exercise intensity and blood lactate is to determine exercise intensity at fixed blood lactate values. This can be done by fitting a regression model that captures the relationship and then "inverse predict" the exercise intensity value. Machado [-Machado_2012] provides a more elaborate method for calculating the lactate threshold. An example of such calculation can be found in [the lecture notes, Linear and curve-linear relationships, and predictions](https://dhammarstrom.github.io/quant-methods).

Your report could use data from the reliability project in the lab. Calculate at least two lactate thresholds (e.g. exercise intensity at 2 and 4 mmol L<sup>-1</sup>) and compare the reliability (typical error as a percentage of the mean) between the two thresholds. If you want to complicate things further you may want to implement other lactate threshold concepts [described in @RN2511; @RN228; Machado_2012] and the course notes.

If you lack data, you could use data from the `exscidata` package or from a previous year (see this file [Lactate threshold tests](/data/reliability-old/laktatprofil-g2.xlsx)).


```{r}
#| code-fold: true
#| message: false
#| warning: false
#| label: "Innlastning av nødvendige pakker og datasett"

library(readxl)
library(ggplot2)
library(magrittr)
library(tidyverse)
library(exscidata)
library(gt)
library(ggtext)
la.profil <- read_xlsx("data/laktatprofil-g2.xlsx")



```

## Selve arbeidsoppgave: Velg to terskelverdier og sammenlign reliabiliteten mellom dem, som typical error, prosent av gjennomsnitt(?)


```{r}
#| code-fold: true
#| message: false
#| warning: false
#| fig-cap: "Figur 1: Viser laktatverdi på 225W og predikert 90% av VO2maks"
#| fig-cap-location: bottom


dat <- cyclingstudy |> 
    filter(timepoint == "pre") |> 
  select(subject, timepoint, group, lac.225, VO2.max, HF.max) |> 
mutate(nittipros_hfmax = HF.max * 0.9)

# Lag funksjon lm

lm_dat <- lm(dat$nittipros_hfmax ~ lac.225, data = dat)


lm_fit <- data.frame(fitted = lm_dat$fitted.values,
                     lm_dat$model,
                     resid = lm_dat$residuals)

lm_fit |> 
ggplot(aes(lac.225, fitted)) +
  geom_point(size = 3, fill = "gold", shape = 22) +
  theme_minimal() +
    geom_point(aes(lac.225, dat$nittipros_hfmax)) +
  geom_segment(aes(x = lac.225, xend = lac.225, y = fitted, yend = fitted + resid)) +
  labs(y = "Fitted values 90% av makshf",
       x = "Laktat @225W")

summary(lm_dat)






```


## Reliabilitetsregning 


```{r}
#| code-fold: true
#| label: "lac2 @2&4 pre"
#| output: FALSE


dat_pre <- cyclingstudy |> 
  select(timepoint, lac.225, VO2.max) |> 
  filter(timepoint == "pre")

lm_dat_lac2_pre <- lm(VO2.max ~ lac.225, data = dat_pre) 


lm_val_lac2_pre <- data.frame(lm_dat_lac2_pre$model,
                     fitted = lm_dat_lac2_pre$fitted.values,
                     resid = lm_dat_lac2_pre$residuals)

exp_vo2_lac2_pre <- coef(lm_dat_lac2_pre) [1] + coef(lm_dat_lac2_pre) [2] * 2

exp_vo2_lac4_pre <- coef(lm_dat_lac2_pre) [1] + coef(lm_dat_lac2_pre) [2] * 4

exp_vo2_lac2_pre

exp_vo2_lac4_pre

```

```{r}
#| code-fold: true
#| output: FALSE
#| label: "lac @2&4 meso3"

dat_m3 <- cyclingstudy |> 
  select(timepoint, lac.225, VO2.max) |> 
  filter(timepoint == "meso3")

lm_dat_lac2_meso3 <- lm(VO2.max ~ lac.225, data = dat_m3) 


lm_val_lac2_meso3 <- data.frame(lm_dat_lac2_meso3$model,
                     fitted = lm_dat_lac2_meso3$fitted.values,
                     resid = lm_dat_lac2_meso3$residuals)

exp_vo2_lac2_meso3 <- coef(lm_dat_lac2_meso3) [1] + coef(lm_dat_lac2_meso3) [2] * 2

exp_vo2_lac4_meso3 <- coef(lm_dat_lac2_meso3) [1] + coef(lm_dat_lac2_meso3) [2] * 4

exp_vo2_lac2_meso3

exp_vo2_lac4_meso3



```


```{r}

exp_df <- tibble(exp_vo2_lac4_pre, exp_vo2_lac4_meso3, exp_vo2_lac2_pre, exp_vo2_lac2_meso3) 

```





